---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- build with rmarkdown::render("README.Rmd") -->

```{r setup, include = FALSE}
library(infx)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# [infx](https://nbenn.github.io/infx)

[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://img.shields.io/badge/lifecycle-maturing-blue.svg)
[![Travis-CI Build Status](https://travis-ci.org/nbenn/infx.svg?branch=master)](https://travis-ci.org/nbenn/infx)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/nbenn/infx?branch=master&svg=true)](https://ci.appveyor.com/project/nbenn/infx)
[![Coverage status](https://codecov.io/gh/nbenn/infx/branch/master/graph/badge.svg)](https://codecov.io/github/nbenn/infx?branch=master)

Access to [InfectX](http://www.infectx.ch)/[TargetInfectX](https://www.targetinfectx.ch) screening data from R. A browser-based view of the data is available [here](http://www.infectx.ch/databrowser).

## Installation

You can install [infx](https://nbenn.github.io/infx) from Github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("nbenn/infx")
```

## InfectX

[InfectX](http://www.infectx.ch) and its successor project [TargetInfectX](https://www.targetinfectx.ch) are large-scale high throughput screening experiments focused on the human infectome of a set of viral and bacterial pathogens. In order to identify host-provided components involved in pathogen entry and host colonization, several RNA interference screens were carried out on HeLa cells, using siRNA libraries from vendors such as Dharmacon, Quiagen and Ambion. Of the many performed screens, currently the data of kinome-wide screens for five bacterial pathogens (*Bartonella henselae*, *Brucella abortus*, *Listeria monocytogenes*, *Salmonella* typhimurium, and *Shigella flexneri*) and three viruses (Adenovirus, Rhinovirus, and *Vaccinia virus*) is publicly available[^1]. Additionally, several genome-wide screens will follow suit in the coming months.

All collected data, including raw imaging data, [CellProfiler](http://cellprofiler.org) derived feature data and infection scoring at single cell resolution, alongside extensive metadata, is hosted by the laboratory information management system [openBIS](https://wiki-bsse.ethz.ch/display/bis/Home). This R package provides access to the openBIS [JSON-RPC API](https://wiki-bsse.ethz.ch/display/openBISDoc1304/openBIS+JSON+API), enabling listing of data organization objects, searching for and downloading of data sets.

## OpenBIS

This document gives only a brief introduction on how to work with openBIS. For a more in-depth information on how to use this package, please refer to the vignettes ["Introduction to infx"](https://nbenn.github.io/infx/articles/infx-intro.html) and ["JSON object handling"](https://nbenn.github.io/infx/articles/json-class.html). For an extensive look at what parts of the API are implemented by this package and how to extend the package to support further functionality, have a look at the vignette ["OpenBIS API coverage"](https://nbenn.github.io/infx/articles/openbis-api.html). Documentation of exported functions is available from within the R help system or [here](https://nbenn.github.io/infx/reference/index.html).

For every API call, a valid login token is required. Tokens can be created using [`login_openbis()`](https://nbenn.github.io/infx/reference/login.html) and tested for validity with [`is_token_valid()`](https://nbenn.github.io/infx/reference/login.html).

```{r login}
tok <- login_openbis(user = "rdgr2014",
                     pwd = "IXPubReview",
                     auto_disconnect = FALSE)
is_token_valid(tok)
```

Using the valid login token, openBIS can now be queried, for example for a list of all projects that are available for the given user, using [`list_projects()`](https://nbenn.github.io/infx/reference/list_projects.html).

```{r projects}
projects <- list_projects(tok)
projects[[1]]
```

Finally, the login token should be destroyed, using [`logout_openbis()`](https://nbenn.github.io/infx/reference/login.html).

```{r logout}
logout_openbis(tok)
is_token_valid(tok)
```

[^1]: [*BMC Genomics* 2014 **15**:1162](https://doi.org/10.1186/1471-2164-15-1162)
