---
title: "OpenBIS API coverage"
author: "Nicolas Bennett"
date: "`r Sys.Date()`"
output:
  html_vignette:
    self_contained: no
vignette: >
  %\VignetteIndexEntry{OpenBIS API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(infx)
library(rvest)
library(knitr)
```

```{r api-calls, include = FALSE}
get_openbis_items <- function(x) {
  sections <- tools:::RdTags(x) == "\\section"
  if (sum(sections) > 0L)
    sections[sections] <- sapply(x[sections], function(sec) {
      tolower(tools:::.Rd_get_text(sec[[1L]])) == "openbis"
    })

  if (sum(sections) == 1L) {
    obis <- x[[which(sections)]][[2L]]
    is_itemize <- tools:::RdTags(obis) == "\\itemize"
    res <- unlist(obis[is_itemize])
    res[!res %in% c("\n", " ")]
  } else if (sum(sections) > 1L)
    stop("expecting one or zero openbis sections")
  else
    NA
}

pkg <- "infx"
rd_db <- tools:::fetchRdDB(file.path(find.package(pkg), "help", pkg))
api_calls <- Filter(Negate(is.na), unlist(lapply(rd_db, get_openbis_items)))

urls <- gsub("^\\\\href\\{", "",
             sapply(strsplit(api_calls, "\\}\\{"), `[`, 1L))

api_calls <- gsub("\\}$", "", sapply(strsplit(api_calls, "\\}\\{"), `[`, 2L))
api_calls <- strsplit(api_calls, ":")
```

```{r api-info, include = FALSE}
info <- lapply(sort(unique(urls)), function(x) {
  docs <- read_html(x)
  api <- html_nodes(docs,
                    xpath = "/html/body/div[4]/div[2]/ul/li/ul[2]/li/table")

  tab <- apply(html_table(api)[[1]], 1, function(y) {
    y <- y[2]
    fun_name <- sub("\\($", "", regmatches(y, regexpr("^.+?\\(", y)))
    desc <- sub("\\)\n", "", regmatches(y, regexpr("\\)\n.+$", y)))

    if (length(desc) && grepl("^Deprecated", desc))
      return(NULL)

    found <- fun_name %in% sapply(api_calls[urls == x], `[`, 2L)

    list(`Method name` = fun_name,
         Status = ifelse(found, "implemented", "skipped"),
         Description = sub("IDssServiceRpcScreening\\.",
                           "IDssServiceRpcScreening ",
                           sub("ch\\..+\\.dto\\.", "",
                               gsub("\n", "",
                                    sub("^/\\*\\* ", "", desc)))))
  })

  title <- html_text(html_nodes(docs, xpath = "/html/body/div[3]/h2"))
  desc <- html_nodes(docs, xpath = "/html/body/div[4]/div[1]/ul/li/div")
  desc <- strsplit(html_text(desc), "\\.")[[1]]
  desc <- paste0(desc[1], ". More information available [here](", x, ").\n")
  tabl <- do.call(rbind, tab)
  link <- paste0("[", sub("^Interface ", "", title), "](", x, ")")

  list(title = title,
       description = desc,
       table = tabl,
       status = round(mean(tabl[, 2] == "implemented") * 100),
       link = link)
})
```

This vignette serves to document which sections of the openBIS API are implemented by the presented client package and which functionality was omitted. Furthermore, the basic mechanisms for making requests to openBIS are explained such that a user who wishes to add some of the omitted API functions has some information on how to extend `infx`. Some general documentation on the API, offered by the openBIS developers is available [here](https://wiki-bsse.ethz.ch/display/openBISDoc1304/openBIS+JSON+API).

## Creating a REST call

The basic functionality powering all REST calls of `infx` is provided by `do_requests_serial()` and `do_requests_parallel()`. Both functions take roughly the same input produce identical results, but differ in how curl is called. The former function makes requests in a sequential fashion, iteratively calling `curl::curl_fetch_memory()` and the latter performs asynchronous requests, using `curl::curl_fetch_multi()` to assemble all requests and `curl::multi_run()` to carry them out in asynchronous parallel fashion.

Much of the behavior of `do_requests_serial()` and `do_requests_parallel()` can be customized using three functions passed as arguments `create_handle`, `check` and `finally`. Additionally, `do_requests_*()` takes a vector of `urls` and an optional object `bodies` which is expected to be of the same length as `urls`. Urls can either be passed as a character vector or a list of unevaluated function calls which will be `base::eval()`ued shortly before being used. Both functions support retrying failed requests up to `n_try` times.

For constructing the requests, both `urls` and `bodies` are iterated together. First, the function passed as `create_handle`, which receives as input the current entry of the `bodies` object is used to construct a curl handle using `curl::create_handle()`. Together with the current url entry, `curl::curl_fetch_*()` is called and the resulting object is passed as first argument to the `check` function, which receives as second argument the current entry of the `bodies` object. The function passed as `check` argument should make sure the request completed successfully and in case of failure return a `simpleError` object, created by `base::simpleError()`. In case of success, it should return the part of the curl response object that is of further interest (most likely the `content` entry). Next the `finally` function is called on the object returned by `check` to do some final processing (e.g. parsing JSON or reading a binary file). If the `check` function signaled a failure and the number of available tries as specified by `n_try` is not used up, the request is made again. If the allowed number of tries is exceeded, a warning is issued.

Asynchronous requests are implemented by first adding `n_con` curl handles to a new multi handle and starting the downloads by calling `curl::multi_run()`. For each successful request, a new handle is added to the pool using the `done` callback function of `curl::multi_add()`. In conjunction with passing urls as unevaluated function calls, this helps with urls that have a limited lifetime in that the url is only created right before being consumed. Instead of adding all requests at the same time and letting curl handle queuing, only `n_con` requests are handled by curl at any given time.

A very basic use of `do_requests_serial()` is shown in the following example. The default function of the `create_handle` argument creates a new clean curl handle with no options set. The default function of the `check` argument makes sure that the status code is equal to 200 and returns the `content` entry of the list returned by `curl::curl_fetch_memory()`. The default function for `finally` is `base::identity()`. In order to receive a human-readable result instead of a raw vector, a function `process_raw()` is created, which converts the raw vector to a character vector yielding a JSON string.

```{r rest-simple}
urls <- "https://httpbin.org/ip"

pretty_json <- function(x)
  jsonlite::prettify(rawToChar(x))

do_requests_serial(urls, finally = pretty_json)
```

Note that a list of length 1 is returned by `do_requests_serial()`. This is because the `do_requests_*()` functions are vectorized over requests and therefore return a list with as many entries as requests. A slightly more involved example for creating a request using `do_requests_serial()` is as follows.

```{r rest-json}
urls <- rep("https://httpbin.org/post", 2)
json <- list(list(a = "foo"),
             list(a = "bar"))

post_handle <- function(x)
  curl::handle_setheaders(
    curl::new_handle(postfields = charToRaw(jsonlite::toJSON(x))),
    "Content-Type" = "application/json"
  )
  
process_json <- function(x) 
  jsonlite::fromJSON(rawToChar(x))$json

res <- do_requests_serial(urls, json,
                          create_handle = post_handle,
                          finally = process_json)
identical(res, json)
```

In order to customize the curl handle, a `create_handle` function is supplied, which receives for each request the corresponding entry of the object passed as `bodies` argument. The `finally` function in this example parses the returned JSON object into a list. Since for POST requests, `httpbin` mirrors the POST data, the initial `json` list is returned.

## Creating a JSON-RPC request

Building on functionality offered by the `do_requests_*()` functions, `make_requests()` helps with constructing JSON-RPC calls. As per the JSON-RPC 2.0 specification, a request object is a JSON string with the following members^[[**JSON-RPC 2.0 Specification** Rev. 2013-01-04. JSON-RPC Working Group](http://www.jsonrpc.org/specification)]:

* `jsonrpc`: A String specifying the version of the JSON-RPC protocol.
* `method`: A String containing the name of the method to be invoked.
* `params`: A Structured value that holds the parameter values to be used during the invocation of the method.
* `id`: An identifier established by the Client.

In order to create a list of such request objects, the arguments `methods`, `params`, `ids` and `version`, which map to `method`, `params`, `id` and `jsonrpc`, respectively, are combined and converted to JSON. OpenBIS API endpoint urls are constructed with the helper function `api_url()` which can receive arguments `api_endpoint` and `host_url`, forwarded by `make_requests()`. The argument `n_con` is used to specify the maximal number of simultaneous connections made to the server.

As `make_requests()` is designed to construct several requests at the time, the arguments `methods` and `params` are vectorized. If any of the passed objects if of length 1, it is `base::rep()`eated to the required length such that all three objects are of the same length. Care has to be taken with the list passed as `params` such that its length corresponds to the number of requests. This means that if only a single set of parameters is passed, this list has to be wrapped by another list.

As a last argument, a `finally` function can be passed, which defaults to `process_json()`. This function first converts all typed JSON objects to `json_class` objects using `as_json_class()`, resolves object references using `resolve_references()` in order to make all `json_class` objects self-contained and then creates a `json_vec` object using `as_json_vec()`, therefore allowing S3 method dispatch on lists of `json_class` objects. For more information on the specifics of this, please refer to the vignette ["JSON object handling"](json-class.html).

For single requests, a wrapper around `make_requests()` is available as `make_request()`. This function wraps the `params` argument in a list such that a list of length 1 is passed to `make_requests()` and returns the first entry of the list resulting from calling `make_requests()`. The following example shows how `make_request()` is used to implement the API method [`listProjects`](https://svnsis.ethz.ch/doc/openbis/13.04.0/ch/systemsx/cisd/openbis/generic/shared/api/v1/IGeneralInformationService.html#listProjects%28java.lang.String%29), which takes an access token passed as a list as its only argument.

```{r simple-rpc}
token <- login_openbis()

projects <- make_request(api_url(api_endpoint = "gis"),
                         method = "listProjects",
                         params = list(token))
print(projects, length = 10L)
```

To illustrate the used of `make_requests()`, the API method [`getDownloadUrlForFileForDataSet`](https://svnsis.ethz.ch/doc/openbis/13.04.0/ch/systemsx/cisd/openbis/dss/generic/shared/api/v1/IDssServiceRpcGeneric.html#getDownloadUrlForFileForDataSet%28java.lang.String,%20java.lang.String,%20java.lang.String%29) is implemented. This function generates a download url for a file in a data set and requires a separate API call for each combination of data set and file path. In order to carry out several of these requests asynchronously, a list has to be passed as `params` argument such that each entry contains a list holding a login token, a dataset code and a file path.

```{r multi-rpc}
dataset_codes <- c("20120629093035782-603380", "20121011143734361-1359915")
file_path <- c("original/aThresholdedInfectionScoring_bBB01-1I.csv")

params <- lapply(dataset_codes, function(x) list(token, x, file_path))

str(params)

donwload_urls <- make_requests(api_url(api_endpoint = "dsrg"),
                               method = "getDownloadUrlForFileForDataSet",
                               params = params)
str(donwload_urls)

logout_openbis(token)
```

In addition to the JSON-RPC calls enabled by `make_request()`/`make_requests()`, the `do_requests_*()` functions are also used for (asynchronous) file downloads in `fetch_files()`. Slightly different functions are used as `create_handle` and `check` arguments but apart from that the logic remains largely the same.

Any arguments passed as `...` to `make_requests()` will be forwarded to `api_url()` which creates an api endpoint url and passes this on to the `do_requests_*()` functions. As all functions that issue API calls use `make_requests()`, this makes possible to not only target the InfectX openBIS instance, but arbitrary openBIS servers that support the v1 JSON-RPC API. The following example serves to illustrate how this mechanism can be used to access the openBIS demo.

```{r third-party, fig.align = "center"}
token <- login_openbis("test_observer", "test_observer",
                       host_url = "https://openbis.elnlims.ch")

gel_data <- search_openbis(
  token,
  search_criteria(
    attribute_clause("type", "ELN_PREVIEW"),
    sub_criteria = search_sub_criteria(
      search_criteria(
        attribute_clause(value = "/DIANA_OTTOZ/WB_LEXA-ER-B112")
      ),
      type = "sample"
    )
  )
)

gel_img <- fetch_files(token, gel_data,
                       reader = magick::image_read)

attributes(gel_img[[1]])
print(gel_img[[1]])

logout_openbis(token)
```

`api_url()` can accept arguments `api_endpoint`, `host_url` and `full_url` and ignores any further arguments. The first argument offers a selection of hard-coded API endpoints and in combination with the second argument is used to access one of them on the specified openBIS host. In case an url that is not supported by the `api_endpoint` selector is desired, the complete url can
be passed as `full_url` which will simply be returned by `api_url()`. 

## Summary of API methods

Currently no methods from the API sections

* [IGeneralInformationChangingService](http://svnsis.ethz.ch/doc/openbis/13.04.0/ch/systemsx/cisd/openbis/generic/shared/api/v1/IGeneralInformationChangingService.html)
* [IQueryApiServer](http://svnsis.ethz.ch/doc/openbis/13.04.0/ch/systemsx/cisd/openbis/plugin/query/shared/api/v1/IQueryApiServer.html)
* [IWebInformationService](http://svnsis.ethz.ch/doc/openbis/13.04.0/ch/systemsx/cisd/openbis/generic/shared/api/v1/IWebInformationService.html)

are implemented as they mostly deal with modifying metadata and creating new aggregation reports. The main focus of this package is retrieving data and therefore only methods from the API sections

```{r api-summary, echo = FALSE, results = "asis", tidy = FALSE}
for (i in seq_along(info))
  cat(paste0("* ", info[[i]]$link, ", ", info[[i]]$status), "% implemented\n")
```

are currently available. A more detailed overview of what functionality is implemented in this API client is given in the following sections.

```{r api-tables, echo = FALSE, results = "asis", tidy = FALSE}
for (i in seq_along(info)) {
  cat(paste0("###", info[[i]]$title, "\n", info[[i]]$description, "\n"))
  print(kable(info[[i]]$table))
  cat("\n")
}
```
